CO 789, Homework 1

### 2024-02-21 co789/homework-03/q3/main.tex
homework 03 draft

\section*{Question 6}
\subsection*{(a)}
We denote a polynomial with degree $\phi(n) - 1$ by $f(x) = f_0 + f_1x + \ldots + f_{\phi(n)-1}x^{\phi(n)-1}$. It is trivially true that if $f(x) = g(x)$, then $f(\zeta_i) = g(\zeta_i)$ for all $1 \leq i \leq \phi(n)$.

On the other hand, if for two polynomials $f, g$ with degree $\phi(n) - 1$, their NTT representations are exactly identical, then for all $1 \leq i \leq \phi(n)$:

$$
f(\zeta_i) = g(\zeta_i)
$$

Define $h(x) = f(x) - g(x)$, then $\zeta_1, \zeta_2, \ldots, \zeta_{\phi(n)}$ are distinct roots of $h(x)$. This means that $h(x)$ must have form:

$$
h(x) = h^\prime(x)(x - \zeta_1)(x - \zeta_2) \ldots (x - \zeta_{\phi(n)})
$$

Because $f(x), g(x)$ both have degree $\phi(n) - 1$, the degree of $f(x) - g(x)$ cannot be more than $\phi(n) - 1$. Therefore, $h(x)$ must be $0$, because otherwise it must have degree of at least $\phi(n)$. $\blacksquare$

\subsection*{(b)}
Recall the NTT transformation described in part (1): $\operatorname{NTT}(f) \mapsto (f(\zeta_1), f(\zeta_2), \ldots, f(\zeta_{\phi(n)}))$. From here we know:

$$
\begin{aligned}
\operatorname{NTT}(f \cdot g) 
&= ((f \cdot g)(\zeta_1), (f \cdot g)(\zeta_2), \ldots, (f \cdot g)(\zeta_{\phi(n)})) \\
&= (f(\zeta_1) \cdot g(\zeta_1), f(\zeta_2) \cdot g(\zeta_2), \ldots, f(\zeta_{\phi(n)}) \cdot g(\zeta_{\phi(n)})) \\
&= (f(\zeta_1), f(\zeta_2), \ldots, f(\zeta_{\phi(n)})) \circ (g(\zeta_1), g(\zeta_2), \ldots, g(\zeta_{\phi(n)})) \\
&= \operatorname{NTT}(f) \circ \operatorname{NTT}(g)
\end{aligned}
$$

$\blacksquare$

>>>>

\section*{Question 3}

Observe the following:

$$
\begin{aligned}
A\mathbf{z} &= A(\mathbf{y} + c\mathbf{s}) \\
&= \mathbf{w} + cA\mathbf{s} \\
&= \mathbf{w} + c(\mathbf{t} - \mathbf{e})
\end{aligned}
$$

Assuming that $c$ is invertible, re-arranging the equation above gives us:

$$
\mathbf{e} = \mathbf{t} - c^{-1}(A\mathbf{z} - \mathbf{w})
$$

From here we can attempt to recover the secret key $\mathbf{s}$ by solving $A\mathbf{s} = \mathbf{t} - \mathbf{e}$. While this is an instance of inhomogeneous SIS, with proto-Dilithium $A$ is a wide matrix, which makes SIS easier to solve. In the lecture notes, we simply assume that wide SIS is "too easy".

### 2024-02-21 co789/homework-03/q5/main.tex
homework 03 draft

\section*{Question 1}

>>>>

\section*{Question 5}

\subsection*{1}
For a given message $m$, we can forge a signature under the public key $(A, \mathbf{t})$ using the following procedure:

\begin{enumerate}
    \item Sample $\mathbf{y} \leftarrow \chi_y$ 
    \item Compute $\mathbf{w} \leftarrow A\mathbf{y}$
    \item Compute $c \leftarrow H(\mathbf{w}, \mathbf{t})$
    \item Give $(A, \mathbf{w} + c\mathbf{t})$ to the $\operatorname{module-ISIS}(k, l, q, p(x), \gamma_1 - np\tau)$ solver, which returns some $\mathbf{z}$
    \item Output $\sigma = (\mathbf{w}, c, \mathbf{z})$
\end{enumerate}

$\sigma$ is a valid forgery because $\mathbf{z}$ as a solution to the $\operatorname{module-ISIS}(k, l, q, p(x), \gamma_1 - np\tau)$ problem satisfies the verification conditions:

\begin{enumerate}
    \item $\norm{\mathbf{z}}_\infty \leq \gamma_1 - np\tau$
    \item $A\mathbf{z} = \mathbf{w} + c\mathbf{t}$
\end{enumerate}

\subsection*{2}
The key recovery attack is as follows:

\begin{enumerate}
    \item Sample some random message $m$
    \item Query the signature of $m$, which is $(\mathbf{w}, c, \mathbf{z})$
    \item Give $(A, \mathbf{w})$ to $\operatorname{module-ISIS}(k, l, q, p(x), \gamma_1)$ solver, which returns some $\mathbf{y}$
    \item Compute $\mathbf{s} = c^{-1}(\mathbf{z} - \mathbf{y})$. $\mathbf{s}$ is the secret key
\end{enumerate}

This procesure works because for the queried signature to be valid, it must satisfy $\mathbf{z} = \mathbf{y} + c\mathbf{s}$, meaning that if we can recover $\mathbf{y}$, then we can recover $\mathbf{s}$. In this instance, a valid $\mathbf{y}$ is recovered using the solver on $A\mathbf{y} = \mathbf{w}$.
